---
title: "covid 12 19 2020"
author: "Dom Pizzano"
date: "12/19/2020"
---
## Total Vaccine Distribution Total for the World 
To track the vaccination efforts, I scrapped vaccination data from the web using publicaly available data sources to then 

Data from: Our World in Data, <https://ourworldindata.org/>.
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=11,fig.height=6}
library(dplyr)
library(ggplot2)
library(usmap)
library(tidyverse)
#data from csv file 
X <- read.csv(url("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv"))

#only getting the most recent totals from the csv data by filtering on the max date
df <- X %>% group_by(location) %>% filter(date == max(date))
```
Location and Date (YYYY-MM-DD) data was last updated:
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,tidy=TRUE}
print(df[1:2])
```

#### Figure 1: Global Covid 19 Vaccination total by Location
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=11,fig.height=6}

#changing the column names in the data frame to be more easily understanable
colnames(df) <- c("Location","Date","Total_Vaccines_Distributed","Total_Vaccinations_Per_Hundred")

#global vaccinations total across the world by country and also the total distributed in the world (most up to date from data source)
ggplot(data=df, aes(x=Location, y=Total_Vaccines_Distributed, fill=Location)) +
  geom_bar(stat="identity") 
```

However this Graph is slightly misleading because it does not put into account total populations of the Locations being observed, when it comes the total vaccinations per 100 people the graph looks like this.

#### Figure 2: Global Covid 19 Vaccination total per 100 people by Location
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=11,fig.height=6}
#global vaccinations by total vaccinations per hunred people in the country
ggplot(data=df, aes(x=Location, y=Total_Vaccinations_Per_Hundred, fill=Location)) + geom_bar(stat="identity")
```

## Total Vaccine (Pfizer) Distribution Total for the United States 
Data from CDC: Center of Disease Control and Prevention, <https://data.cdc.gov/Vaccinations/COVID-19-Vaccine-Initial-Allocations-Pfizer/saz5-9hgg/data>.

#### Figrue 3: Top 10 as Percentage of the Population (2015) Vaccinated with Covid-19 Vaccine
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=8,fig.height=6}

#read in the united states covid 19 vaccination data by state (aka Jurisdiction)
statesDf <- read.csv(url("https://data.cdc.gov/api/views/saz5-9hgg/rows.csv?accessType=DOWNLOAD"))
#must scarp out the special character *** in data
statesDf$Jurisdiction <- str_replace_all(statesDf$Jurisdiction, "[[:punct:]]", "")
#get rid of comma in doses column
statesDf$Total.Pfizer.Allocation..First.Dose..Shipments <- str_replace_all(statesDf$Total.Pfizer.Allocation..First.Dose..Shipments, "[[:punct:]]", "") %>% strtoi()
#change column names to perform a full join, population data is from 2015
colnames(statepop) <- c("fips","abr","Jurisdiction","pop_2015")

#inner join to include
final_df <- inner_join(statesDf,statepop,by="Jurisdiction")  %>% mutate(percent_vaccinated = (Total.Pfizer.Allocation..First.Dose..Shipments/pop_2015)*100)

#getting top ten most vaccinated states by percent vaccinated of population
topTenStatesVP <- head(arrange(final_df,desc(percent_vaccinated)), n = 10)

#getting top ten most vaccinated states by total number of vaccinations
topTenStatesP <- head(arrange(final_df,desc(Total.Pfizer.Allocation..First.Dose..Shipments)), n = 10)

#top ten vaccinated states by percentage of population
ggplot(data=topTenStatesVP, aes(x = reorder(abr, sort(percent_vaccinated,TRUE)), y=percent_vaccinated, fill=abr)) +
  geom_bar(stat="identity")

```

#### Figrue 4: Top 10 as Total Vaccinated with Covid-19 Vaccine
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=8,fig.height=6}
#top ten vaccinated states by total
ggplot(data=topTenStatesP, aes( x = reorder(abr, sort(Total.Pfizer.Allocation..First.Dose..Shipments,TRUE)), y = Total.Pfizer.Allocation..First.Dose..Shipments, fill=abr)) +
  geom_bar(stat="identity")
```

#### Figure 5: United States Density Map of Coverage Percentage of Population (2015) for Vaccines Distributed (Pfizer)
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}
#us map of states with highest percentage of people vaccinated 
 plot_usmap(data = final_df, values = "percent_vaccinated", color = "black") + 
   scale_fill_continuous(
     low = "white", high = "red", name = "Percent Vaccinated", label = scales::comma
   ) + theme(legend.position = "right")
```

#### Figure 6: United States Density Map of Total Vaccinations Distributed (Pfizer)
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}
# total vaccines distributed per state
 plot_usmap(data = final_df, values = "Total.Pfizer.Allocation..First.Dose..Shipments", color = "black") + 
   scale_fill_continuous(
     low = "white", high = "blue", name = "Total Vaccinated", label = scales::comma
   ) + theme(legend.position = "right")

```

This report is meant to give insight into what states and countries were able to vaccinate the most people during the initail days of availablity for Vaccines, only Pfizer vaccine data was available for this week in the United States. Since the data comes from websites that are being updated (hopefully), next week if I re-run this R report, their should be a new figures and updated data. 

