---
title: "Covid19 Vaccinations 2-23-2021"
author: "CodeByDom"
date: "2/23/2021"
knit: (function(input_file, encoding) {
  out_dir <- './';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, '2-23-2021.html'))})
---
## Total Vaccine Distribution Total for the World 
To track the vaccination efforts, I scrapped vaccination data from the web using publicly available data sources. 

Data from: Our World in Data, <https://ourworldindata.org/>.
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=11,fig.height=6}
library(dplyr)
library(ggplot2)
library(usmap)
library(tidyverse)
#data from csv file 
X <- read.csv(url("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv"))

#only getting the most recent totals from the csv data by filtering on the max date
df <- X %>% group_by(location) %>% filter(date == max(date))
```
Location and Date (YYYY-MM-DD) data was last updated:
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,tidy=TRUE}
print(df[1:3])
```

#### Figure 1: Global Covid 19 Vaccination Total by Location (Top 15)
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=20,fig.height=14}
df15 <- head(arrange(df,desc(total_vaccinations)), n = 15)
ggplot(data=df15, aes(x=reorder(location, sort(total_vaccinations,FALSE)), y=total_vaccinations, fill=location)) +
  geom_bar(stat="identity") + xlab("Countries") + ylab("Total Vaccinations Given")+theme(axis.text=element_text(size=12),
        axis.title=element_text(size=18,face="bold"))
```

However, this graph is slightly misleading because it does not put into account the total populations of the Locations being observed, when it comes to the total vaccinations per 100 people the graph looks like this.

#### Figure 2: Global Covid 19 Vaccination total per 100 people by Location (Top 15)
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=20,fig.height=14}
df152 <- head(arrange(df,desc(total_vaccinations_per_hundred)), n = 15)
ggplot(data=df152, aes(x=reorder(location, sort(total_vaccinations_per_hundred,TRUE)), y=total_vaccinations_per_hundred, fill=location)) +
  geom_bar(stat="identity") + xlab("Countries") + ylab("Total Vaccinations Given per 100 People")+theme(axis.text=element_text(size=12),
        axis.title=element_text(size=18,face="bold"))
```

#### Figure 1.2: Total Daily Vaccinations by Location (Top 15) 
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=20,fig.height=14}
df153 <- head(arrange(df,desc(daily_vaccinations)), n = 15)
ggplot(data=df153, aes(x=reorder(location, sort(daily_vaccinations,FALSE)), y=daily_vaccinations, fill=location)) +
  geom_bar(stat="identity") + xlab("Countries") + ylab("Total Vaccinations Given Daily")+theme(axis.text=element_text(size=12),
        axis.title=element_text(size=18,face="bold"))
```

#### Figure 2.2: Total Daily Vaccinations per Million by Location (Top 15) 
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=20,fig.height=14}
df154 <- head(arrange(df,desc(daily_vaccinations_per_million)), n = 15)
ggplot(data=df154, aes(x=reorder(location, sort(daily_vaccinations_per_million,FALSE)), y=daily_vaccinations_per_million, fill=location)) +
  geom_bar(stat="identity") + xlab("Countries") + ylab("Total Vaccinations Given Daily per Million")+theme(axis.text=element_text(size=12),
        axis.title=element_text(size=18,face="bold"))
```

## Total Vaccine (Pfizer) Distribution Total for the United States 
Data from CDC: Center of Disease Control and Prevention, <https://data.cdc.gov/Vaccinations/COVID-19-Vaccine-Initial-Allocations-Pfizer/saz5-9hgg/data>.

#### Figrue 3: Top 10 as Percentage of the Population (2015) Vaccinated with Covid-19 Vaccine (Pfizer)
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=8,fig.height=6}

#read in the united states covid 19 vaccination data by state (aka Jurisdiction)
statesDf <- read.csv(url("https://data.cdc.gov/api/views/saz5-9hgg/rows.csv?accessType=DOWNLOAD"))
#must scarp out the special character *** in data
statesDf$Jurisdiction <- str_replace_all(statesDf$Jurisdiction, "[[:punct:]]", "")
#get rid of comma in doses column
statesDf$Total.Pfizer.Allocation..First.Dose..Shipments <- str_replace_all(statesDf$Total.Pfizer.Allocation..First.Dose..Shipments, "[[:punct:]]", "") %>% strtoi()
#change column names to perform a full join, population data is from 2015
colnames(statepop) <- c("fips","abr","Jurisdiction","pop_2015")

#inner join to include
final_df <- inner_join(statesDf,statepop,by="Jurisdiction")  %>% mutate(percent_vaccinated = (Total.Pfizer.Allocation..First.Dose..Shipments/pop_2015)*100)

#getting top ten most vaccinated states by percent vaccinated of population
topTenStatesVP <- head(arrange(final_df,desc(percent_vaccinated)), n = 10)

#getting top ten most vaccinated states by total number of vaccinations
topTenStatesP <- head(arrange(final_df,desc(Total.Pfizer.Allocation..First.Dose..Shipments)), n = 10)

#top ten vaccinated states by percentage of population
ggplot(data=topTenStatesVP, aes(x = reorder(abr, -percent_vaccinated), y=percent_vaccinated, fill=abr)) +
  geom_bar(stat="identity") + xlab("States") + ylab("Percent of Population Covered By Distributed Vaccines")

```

#### Figrue 4: Top 10 as Total Vaccinated with Covid-19 Vaccine (Pfizer)

```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=8,fig.height=6}
#top ten vaccinated states by total
ggplot(data=topTenStatesP, aes( x = reorder(abr, -Total.Pfizer.Allocation..First.Dose..Shipments), y = Total.Pfizer.Allocation..First.Dose..Shipments, fill=abr)) +
  geom_bar(stat="identity")+ xlab("States") + ylab("Total Number of Vaccines")
```

#### Figure 5: United States Density Map of Coverage Percentage of Population (2015) for Vaccines Distributed (Pfizer)
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}
#us map of states with highest percentage of people vaccinated 
 plot_usmap(data = final_df, values = "percent_vaccinated", color = "black") + 
   scale_fill_continuous(
     low = "white", high = "red", name = "Percent Vaccinated", label = scales::comma
   ) + theme(legend.position = "right")
```

#### Figure 6: United States Density Map of Total Vaccinations Distributed (Pfizer)
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}
# total vaccines distributed per state
 plot_usmap(data = final_df, values = "Total.Pfizer.Allocation..First.Dose..Shipments", color = "black") + 
   scale_fill_continuous(
     low = "white", high = "blue", name = "Total Allocated", label = scales::comma
   ) + theme(legend.position = "right")

```


## Total Vaccine (Moderna) Distribution Total for the United States 

#### Figure 7: Top Ten States with Vaccine (Moderna) Distributed as Percent of Population (2015) across the United States
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=14,fig.height=6}
modDf <- read.csv(url("https://data.cdc.gov/api/views/b7pe-5nws/rows.csv?accessType=DOWNLOAD"))
 
 
#must scarp out the special character *** in data
modDf$Jurisdiction <- str_replace_all(statesDf$Jurisdiction, "[[:punct:]]", "")

#get rid of comma in doses column
modDf$Total.Moderna.Allocation..First.Dose..Shipments <- str_replace_all(modDf$Total.Moderna.Allocation..First.Dose..Shipments, "[[:punct:]]", "") %>% strtoi()

final_df2 <- inner_join(modDf,statepop,by="Jurisdiction")  %>% mutate(percent_vaccinated = (Total.Moderna.Allocation..First.Dose..Shipments/pop_2015)*100)
topTenStatesmVP <- head(arrange(final_df2,desc(percent_vaccinated)), n = 10)

#getting top ten most vaccinated states by total number of vaccinations
topTenStatesmP <- head(arrange(final_df2,desc(Total.Moderna.Allocation..First.Dose..Shipments)), n = 10)

#top ten vaccinated states by percentage of population

ggplot(data=topTenStatesVP, aes(x = reorder(Jurisdiction,-percent_vaccinated), y=percent_vaccinated, fill=Jurisdiction)) +
  geom_bar(stat="identity")+ xlab("States") + ylab("Percent of People Covered By Distributed Vaccine")

```

#### Figure 8: Top Ten States with Allocated Vaccines (Moderna)

```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=14,fig.height=6}
#top ten vaccinated states by total
ggplot(data=topTenStatesmP, aes( x = reorder(Jurisdiction,-Total.Moderna.Allocation..First.Dose..Shipments), y = Total.Moderna.Allocation..First.Dose..Shipments, fill=Jurisdiction)) +
  geom_bar(stat="identity") + xlab("States") + ylab("Total Number of Vaccines")

```

#### Figure 9: Density map of the Percent Covered with Moderna Vaccinations 

```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}
plot_usmap(data = final_df2, values = "percent_vaccinated", color = "black") + 
  scale_fill_continuous(
    low = "white", high = "orange", name = "Percent Vaccinated", label = scales::comma
  ) + theme(legend.position = "right")

```

#### Figure 10: Density map of the Total Vaccines (Moderna) Allocated
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}
# total vaccines distributed per state
plot_usmap(data = final_df2, values = "Total.Moderna.Allocation..First.Dose..Shipments", color = "black") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Total Allocated", label = scales::comma
  ) + theme(legend.position = "right")

```

## Total Vaccine (Pfizer + Moderna) Distribution Total for the United States 

#### Figure 11: Top Ten States with Allocated Vaccines as Percentage of Population (Pfizer + Moderna)

```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}
final_df2$pop_2015 <- NULL
final_df2$fips <- NULL
finalDF3 <- inner_join(final_df,final_df2,by="Jurisdiction")  %>% mutate(total_vaccinated = Total.Moderna.Allocation..First.Dose..Shipments+Total.Pfizer.Allocation..First.Dose..Shipments,total_per = total_vaccinated/pop_2015*100)
topTenStatesTotP <- head(arrange(finalDF3,desc(total_per)), n = 10)
topTenStatesTot <- head(arrange(finalDF3,desc(total_vaccinated)), n = 10)
ggplot(data=topTenStatesTotP, aes( x = reorder(Jurisdiction,-total_per), y = total_per, fill=Jurisdiction)) +
  geom_bar(stat="identity") + xlab("States") + ylab("Total Number of Vaccines Distributed as Percent of Population")

```

#### Figure 12: Top Ten States for Total Allocated Vaccines  (Pfizer + Moderna)

```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}
ggplot(data=topTenStatesTot, aes( x = reorder(Jurisdiction,-total_vaccinated), y = total_vaccinated, fill=Jurisdiction)) +
  geom_bar(stat="identity") + xlab("States") + ylab("Total Number of Vaccines Distributed")
```



#### Figrue 13: Total Vaccines (Pfizer + Moderna) Allocated Density Map
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}

# total moderna and pzier vaccines 


plot_usmap(data = finalDF3, values = "total_vaccinated", color = "black") + 
  scale_fill_continuous(
    low = "white", high = "green", name = "Total Vaccinations distributed (Pfizer+Moderna)", label = scales::comma
  ) + theme(legend.position = "right")

```

#### Firgure 14: Total Vaccines (Pfizer + Moderna) Allocated as Percentage of the Population
```{r echo=FALSE,comment = "", prompt=FALSE,message=FALSE,warning=FALSE,fig.width=10,fig.height=6}
plot_usmap(data = finalDF3, values = "total_per", color = "black") + 
  scale_fill_continuous(
    low = "white", high = "purple", name = "Percent of Vaccines covering Population (Pfizer+Moderna)", label = scales::comma
  ) + theme(legend.position = "right")
```

This report is meant to give insight into which states and countries can allocate vaccines to the most people during the initial days of availability for vaccines. Since the data comes from websites with data sources that are being continously updated, next time re-running this R report, there will be new figures and updated data. 

